Using Planning to Generate Natural-Language Sentences:\\ Experiences and Challenges

### 2009-12-16 trunk/Papers/crisp-ff-10/crisp-ff.tex
ff sec ueberarbeitung teil 1

@@

First, while pre-processing the ADL formulas, FF has a sub-routine
that checks the instantiated formulas -- quantifiers already compiled
into grounded conjunctions and disjunctions -- for tautologies. This
involves a loop checking all pairs of sub-formulas within every
conjunction and disjunction, testing whether they are
identical/identical modulo their sign. For example, if we find $\phi$
and $\neg \phi$ within a disjunction, then the disjunction is replaced
by TRUE. Now, we haven't tested thoroughly whether this sub-routine
has any notable effect on the formulas in other benchmarks. It
certainly has no such effect in our domain. The effect it {\em does}
have is take a huge amount of runtime on the long conjunctions that
characterize {\bf \dots} For example {\bf \dots before/after}.

Second, FF's pre-process contains a very small imprudence that
apparently didn't surface in other domains, but that does in ours. FF
distinguishes operators into ``easy'' (DNF precondition) and ``hard''
(non-DNF precondition). The easy ones are instantiated in a
spezialized very effectively implemented procedure, while the hard
ones go through a much less effectively implemented more general
procedure. In particular, that latter procedure uses a data structure
encompassing an integer for every possible instantiation of all
predicates. Since this does not take into account reachability, it is
prohibitively huge in our domain. Now, there actually are no hard
operators in our domain. But FF builds the data structure anyway, even
though it is not needed. For example {\bf \dots before/after.}

@@

@@

While the goal agenda meltdown may be more amusing than alarming, we
next report that the relaxed plan heuristic is not far from such a
meltdown either. They are fatally confused by the weird lifecycle of
the ``subst'' and ``distractor'' atoms: in valid plans, these start
off being false, then they get made true at some point, and then they
are made false again in order to achieve the goal. In other words,
from the perspective of the relaxed plan heuristic, from one state to
the next whole classes of goals suddenly appear ``out of the
blue''. These goals are not accounted for in the preceding state, and
hence the relaxed plan is completely un-informative.

To give just one concrete example, say we are in the initial state,
and have not yet committed to a sentence. The relaxed plan has length
$1$ because the we only need to choose a sentence communicating the
desired meaning. However, in the state $s$ resulting from this action,
the mechanisms of the planning task introduce all the sub-goals that
are required in order to pin-point that meaning (in particular we get
a new goal for all the distractors that must be removed). Hence the
relaxed plan is much longer than before -- arbitrarily longer, in
general. For FF's particular search algorithm, this means that {\em
  enforced hill-climbing turns into a breadth-first search}: it starts
at $h$ value $1$ and thereafter needs to solve the whole problem in
order to find a (goal) state with lower $h$ value.

@@

More generally, from this example we immediately see that, in the
terms of Hoffmann \shortcite{hoffmann:jair-05}, the exit distance from
local minima under $h^+$ is unbounded. Also, we have unrecognized
dead-ends, i.e., dead-end states that have a relaxed plan. An example
is the initial state of a task that features distractors for which
there is no means of removal -- ignoring the distractors, the initial
$h$ value will still be $1$. So our domain is in the most difficult
class of Hoffmann's \shortcite{hoffmann:jair-05} ``plannin domain
taxonomy''. While it shares this property with several other
challenging domains, the extremity of turning enforced hill-climbing
into a breadth-first search is unheard of.

@@

The only FF technique that appears to work reasonably well is its
action pruning technique, the ``helpful actions''. While the relaxed
plan is often much too short, it appears to always contain at least
one of the actions that actually are useful, and it certainly prunes
the set of avaibale actions quite effectively. {\bf \dots now
  something on example and BFS with helpful actions vs. without \dots}

>>>>

%../ff -o xtag-3-2-domain.lisp -f xtag-3-2-problem.lisp -T -H: 0.10
%../ff -o xtag-3-2-domain.lisp -f xtag-3-2-problem.lisp -T: 1.54
%../ff -o xtag-3-2-domain.lisp -f xtag-3-2-problem.lisp -H: 0.30
%../ff -o xtag-3-2-domain.lisp -f xtag-3-2-problem.lisp: 1.80

%../ff -o xtag-5-2-domain.lisp -f xtag-5-2-problem.lisp -T -H: 0.29
%../ff -o xtag-5-2-domain.lisp -f xtag-5-2-problem.lisp -T: seg fault ie too much memory
%../ff -o xtag-5-2-domain.lisp -f xtag-5-2-problem.lisp -H: 11.08
%../ff -o xtag-5-2-domain.lisp -f xtag-5-2-problem.lisp: seg fault ie too much memory

%../ff -o xtag-3-3-domain.lisp -f xtag-3-3-problem.lisp -T -H: 
%../ff -o xtag-3-3-domain.lisp -f xtag-3-3-problem.lisp -T: seg fault ie too much memory
%../ff -o xtag-3-3-domain.lisp -f xtag-3-3-problem.lisp -H: 
%../ff -o xtag-3-3-domain.lisp -f xtag-3-3-problem.lisp: seg fault ie too much memory

@@

First, while pre-processing the ADL formulas, FF has a sub-routine
that checks the instantiated formulas -- quantifiers already compiled
into grounded conjunctions and disjunctions -- for tautologies. This
involves a loop checking all pairs of sub-formulas within every
conjunction and disjunction, testing whether they are
identical/identical modulo their sign. For example, if we find $\phi$
and $\neg \phi$ within a disjunction, then the disjunction is replaced
by TRUE. Now, we haven't tested thoroughly whether this sub-routine
has any notable effect on the formulas in other benchmarks. It
certainly has no such effect in our domain. The effect it {\em does}
have is take a huge amount of runtime on the long goal conjunctions
stating that we do not wish to have any open substitution nodes nor
any distractors. The simple fix is hence to switch this sub-routine
off. In smaller examples, this gives advantages like $0.29s$
vs.\ $11s$ pre-processing time; in a larger example we ran, original
FF's pre-process doesn't finish within 15 minutes whereas our fix
yields a pre-processing time of about $3s$.

Second, FF's pre-process contains a very small imprudence that
apparently didn't surface in other domains, but that does in ours. FF
distinguishes operators into ``easy'' (DNF precondition) and ``hard''
(non-DNF precondition). The easy ones are instantiated in a
spezialized very effectively implemented procedure, while the hard
ones go through a much less effectively implemented more general
procedure. In particular, that latter procedure uses a data structure
encompassing an integer for every possible instantiation of all
predicates. Since this does not take into account reachability, it is
prohibitively huge in our domain. Now, there actually are no hard
operators in our domain. But FF builds the data structure anyway, even
though it is not needed. In small examples, this results in a couple
of seconds runtime overhead. In larger examples, it exhausts memory
and causes FF to terminate with a segmentation fault.

@@

%EXAMPLE missing

@@

While the goal agenda meltdown may be more amusing than alarming, we
next report that relaxed plan heuristics are not far from such a
meltdown either. They are fatally confused by the weird lifecycle of
facts encoding substitution nodes and distractors. In instances where
these constructs are important for the construction of a valid
sentence, these facts are required to be false in the goal, and most
of them are initially false. However, any valid plan must make many of
them true at some intermediate time point (e.g.\ distractors arise
depending on the decisions taken by the planner). Hence, from the
perspective of the relaxed plan heuristic, from one state to the next
whole classes of goals suddenly appear ``out of the blue''. These
goals are not accounted for in the preceding state, and hence the
relaxed plan is completely un-informative.

To give just one concrete example, say we are in the initial state,
and have not yet committed to a sentence. For certain kinds of
sentences ({\bf more concrete here???}), the relaxed plan has length
$1$ because the we only need to choose the desired meaning. However,
in the state $s$ resulting from this action, the mechanisms of the
planning task introduce all the sub-goals that are required in order
to pin-point that meaning (in particular we get a new goal for all the
distractors that must be removed). Hence the relaxed plan is much
longer than before -- arbitrarily longer, in general. For FF's
particular search algorithm, enforced hill-climbing (EHC), this means
that {\em FF performs a blind breadth-first search}: EHC starts at $h$
value $1$ and thereafter needs to solve the whole problem in order to
find a (goal) state with lower $h$ value.

%../ff -o modifiers-10-10-domain.lisp -f modifiers-10-10-problem.lisp -T -H

@@

More generally, from this example we immediately see that, in the
terms of Hoffmann \shortcite{hoffmann:jair-05}, the exit distance from
local minima under $h^+$ is unbounded. Also, we have unrecognized
dead-ends, i.e., dead-end states that have a relaxed plan. An example
is the initial state of a task that features distractors for which
there is no means of removal -- ignoring the distractors, the initial
$h$ value will still be $1$. So our domain is in the most difficult
class of Hoffmann's \shortcite{hoffmann:jair-05} ``planning domain
taxonomy''. While it shares this property with several other
challenging domains, the extremity of turning EHC into a breadth-first
search is unheard of.

@@

The only FF technique that appears to work reasonably well is its
action pruning technique, the ``helpful actions''. While the relaxed
plan is often much too short, it appears to always contain at least
one of the actions that actually are useful, and it certainly prunes
the set of avaibale actions quite effectively. For example, in one of
the instances where EHC turns into a breadth-first search, $16875$
nodes are explored without helpful actions; with helpful actions, this
reduces to $3078$. For larger examples, the factor between the two
numbers tends to become larger. {\bf use here larger version of
  modifiers-10-10???}

Since EHC is often completely inadequate, and helpful actions are
quite effective, our FF configuration directly runs best-first search
with helpful actions (as opposed to original FF which switches to
best-first with{\em out} helpful actions if EHC fails).
